### 2020-01-18 nlp_models.tex
updating text

In the past two years, nearly 100 open source, pre-trained DNNs for Natural Language Processing (NLP) have emerged,
based on the revolutionary Transformer architecture, including variants of BERT, Transformer-XML, GPT, etc.
The Transformer architectures consist of blocks of Attention layers, containing of 2 large, Feed Forward (Linear)
weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
the Attention matrices are significantly larger, and frequently under-correlated, with generally larger
Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).

\paragraph{OpenAI GPT Models}
Here, we use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
the opportunity to analyze the effect of both training the same model with different size data sets,
and increasing sizes of both the data set and architectures.
These models have generated significant media attention because of their remarkable ability to
 generate fake text that appears to be real and the potential misuse of this.
For this reason, the original GPT model was trained on on a deficient data set, rendering
the model interesting but not fully functional.  Later, the same GPT model 
The original GPT and later GPT model have the same architecture and number of layers, but

 however, a much improved GPT2 model has been released, which is remarkably good at creating 
auto-genetated text.

>>>>

In the past two years, nearly 100 open source, pre-trained DNNs for Natural Language Processing (NLP) have emerged,
based on the revolutionary Transformer architecture, including variants of BERT, Transformer-XML, GPT, etc.
The Transformer architectures consist of blocks of Attention layers, containing of 2 large, Feed Forward (Linear)
weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
the Attention matrices are significantly larger, and frequently under-correlated, with generally larger
Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).  Here, we breifly look at a few
popular pretrained NLP DNNs to demonstrate that the Theory of Heavy Tailed Self-Regularization also
applies to NLP models and not just CV models, and to highlight how to use the theory to identify
poorly trained models where the empirical norm metrics will perform poorly.

\paragraph{OpenAI GPT Models}
We use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
the opportunity to analyze the effect of both training the same model with different size data sets,
and increasing sizes of both the data set and architectures.
These models have generated significant media attention because of their remarkable ability to
 generate fake text that appears to be real and the potential misuse of this.
For this reason, the original GPT model was trained on on a deficient data set, rendering
the model interesting but not fully functional.  Later, OpenAI released a much improved model--
GPT2 (small)--which has the same architecture and number of layers as GPT, but has
been trained on a larger and betterd data set (and with other changes), making it
remarkably good at generating (near) human-quality fake text.  
By comparing GPT and GPT2, we can indentify empirical markers indicating that
a model has been poorly trained and may perform poorly when deployed.
